### 2013-10-11 JournalToSubmit2013/methodologiesForNFR.tex

%(aqui propuestas relacionadas, analizar sus caracteristicas, si son MDD, si parte de alto nivel, etc.)