### 2013-10-19 SOFSEM2014/FinalVersion/conclusion.tex
no message

This paper presented the $\pi$SOD-M software method for specifying and designing service based applications in the presence of  non-functional constraints. 
Our proposal enhances the SOD-M method with constraints, policies and contracts  to consider non-functional constraints of applications.
We implemented the proposed meta-models on the Eclipse platform and we illustrated the approach by developping a simple application.

$\pi$SOD-M is being used in an academic environment.
So far, the preliminary results indicate that $\pi$SOD-M approach is useful for the development of complex web service applications. 
We are now working on the definition of a PCM-level meta-model to generate BPEL programs (instead of $\pi$-PEWS).