### 2013-10-22 SOFSEM2014/FinalVersion/proofOfConcept.tex

Consider a tracking crime application where civilians and police share information about criminality in given zones of a city. 
Civilian users signal crimes using Twitter.
Police officers can notify crimes, as well as update information about solving cases.
Some of these information are confidential while other can be shared
to the community of users using this application. 
Users can track crimes in given zones. 
Crime information stored by the system may be visualized on a map. 
Some users have different access rights than others.
For example, police officers have more access rights than civilians.

In order to provide these functionalities, the application uses pre-existing services to provide, store and visualize  information.
The business process defines the logic of the application and is specified in terms of tasks. 
Tasks can be performed by people or computers.

The business process and requirements specifications presented in Figure~\ref{fig:constraints} are instances of the Computation-Independent models of Figure~\ref{fig:piSOD-M}.
The business process is represented as a graph while requirements are given as text boxes.
\begin{figure}[ht!]
\centering
%\subfloat[\textit{Tracking Crimes} - Illustration of the Problem.]
%{\label{fig:problem}\includegraphics[width=0.95\textwidth]{figs/problem}}
%~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. (or a blank line to force the subfig onto a new line)
%\\
%\subfloat[Contraints.]
{\label{fig:trans06}\includegraphics[width=0.95\textwidth]{figs/constraints}}
~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. (or a blank line to force the subfig onto a new line)
\caption{Business process for the tracking crime example.}
\label{fig:constraints}
\end{figure}

In our example, crime processing can start with one of two tasks: \textit{(i) notify a crime}, or \textit{(ii) track a crime}.
Notified crimes are stored in a database. 
Tracked crimes are visualized on a map.
The used can ask for detailed information. 
The application is built upon four services: \textsf{Twitter} and an \textsf{ad-hoc police service}, for notifying crimes; \textsf{Amazon}, used as  persistence service; and  \textsf{Google Maps}, for locating and displaying crimes on a map.

Non-functional requirements are specified by rules and conditions to be verified during the execution of tasks.
In our example we have the following non-functional requirements:
\begin{itemize}
\item Twitter requires authentication and allows three login failures before blocking. 
\item Crime notification needs privileged access.
\item Civilian users can only track crimes for which they have clearance: civilian population cannot track all the crimes notified by the police. 
\item If \textsf{Google Maps} is unavailable, the results are delivered as text. 
\item Querying about crimes without having proper clearance yields an empty map.
\item Access rights to detailed information depends on user clearance and zone assignment for police officers. 
\item The application maintains a detailed log. 
\end{itemize}
The idea about these requirements is to leave the application logic expressed by functional requirements as independent as possible from exceptional situations (like the unavailability of a service) and from conditions under which services are called (for instance, through an authentication protocol). These requirements can be  weaved as activities and branches of the composition or implemented apart. The second option is better because makes the maintenance and evolution of applications easier. For instance, the services called by the application are not hard coded (\textsf{Twitter} and \textsf{Google Maps} in the example), neither the actions to deal with exceptions (replacing another map service or doing nothing).

All the system restrictions are modelled as constraints in this example. 
$\pi$SOD-M provides three types of constraints:
\textit{value}, \textit{business} and \textit{exceptions behaviour} constraints.
Each use case (model) can be associated to one or more constraints\footnote{For a more comprehensive account of $\pi$SOD-M the reader can refer to~\cite{PlacidoThesis2012}.}.
\paragraph*{\textit{$\pi$-UseCase} model:} 
our example has five use cases which represent the
 functions (tasks) and constraints of the system (Figure~\ref{fig:piUC}). 
We do not detail the functional part of the specification, due to lack of space.
The constraints defined are: 
\begin{itemize}
  \item The \textsf{Notify crime} task requires that the user is logged in. 
  This is an example of a \textit{value constraint}, where the value associated to the condition depends on the semantics of the application.
  In this case, it represents the maximum number of allowed login attempts;
\item The \textsf{Store crimes} task requires the verification of the user's clearance (also a value constraint). 
\item In order to perform the \textsf{Track crimes} task,
the contact list of the requesting user must include the user that notified the crime.
This is an example of \textit{business constraint}.
Additionally the requesting user must be logged in.
\item  For the \textsf{View crimes' map} task, the specification defines that if the service \textsf{Google Maps}  is not available, the result is presented as text. This is an example of \textit{exceptional behaviour constraint}. 
The availability of the \textsf{Google Maps} service is verified by a \textit{business constraint}.
 \item The \textsf{Show crimes' details} task is specified to have three constraints: A \textit{value constraint} is defined to verify the user's clearance level; a \textit{business constraint} is used to ensure that the user's clearance is valid for the geographic zone of the crime; another \textit{value constraint} defines that the log is to be maintained.
\end{itemize}
\begin{figure}[t]
\centering
\includegraphics[width=0.8\textwidth]{figs/piUseCase}
\caption{$\pi$-UseCase Model}
\label{fig:piUC}
\end{figure}

A model-to-model transformation is defined to refine the \textit{$\pi$-UseCase} model into the
corresponding \textit{$\pi$-ServiceProcess} model, a more detailed representation. 
This (semi-automatic) transformation  is supported by a tool (described in~\cite{PlacidoThesis2012}).

\paragraph*{\textit{$\pi$-ServiceProcess} model:}
in this model task nodes of the $\pi$-UseCase model are better detailed, by refining the control and data flows, and its constraints are transformed into \textit{contracts} (pre- and post-conditions).
The $\pi$-ServiceProcess model describes the activities of the application and defines contracts for each activity or parts of the application (Figure~\ref{fig:piSP}). The main transformations are: 
\begin{itemize}
\item Tasks of the previous model are transformed into \textit{actions};
\item Actions are grouped into \textit{activities} (in accordance to the business logic).
\item Constraints of the \textit{$\pi$-UseCase} model are transformed into assertions.
\end{itemize}

\begin{figure}[t]
\centering
\includegraphics[width=1.05\textwidth]{figs/piServiceProcess}
\caption{$\pi$-\textit{ServiceProcess} Model}
\label{fig:piSP}
\end{figure}

\paragraph*{\textit{$\pi$-ServiceComposition} model:} 
this model refines the previous model by using the activities to produce the workflow of the application.
The model serves to identify those entities that collaborate with the service process by providing services to execute actions. 
This model identifies the services and functions that correspond to each action in the business process.

\begin{figure}[t]
\centering
\subfloat[$\pi$-ServiceComposition Model]
{\label{fig:piSC}\includegraphics[width=0.95\textwidth]{figs/piServiceComposition}}
~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. (or a blank line to force the subfig onto a new line)
\\[-1mm]
\subfloat[$\pi$-ServiceComposition Policies]
{\label{fig:crimePolicies}\includegraphics[width=0.85\textwidth]{figs/policies}}
~ %add desired spacing between images, e. g. ~, \quad, \qquad etc. (or a blank line to force the subfig onto a new line)
\caption{Service Composition and Policies.}
\label{fig:policies}
\end{figure}
%This model describes the service compositions and their related business
%collaborators (external services), policies, rules associated with a policy and
%the whole application process and functionalities.
%The workflow execution will consider the composition of business services and
%their specific Business Collaborators. 
%Each action will trigger the execution of one or more external services. 
%Before and after each service call, policy rules are verified.

In the case of our crime tracking example, the model produced from the $\pi$-\textit{ServiceProcess} model of Figure~\ref{fig:piSP} is given in Figures~\ref{fig:piSC} and~\ref{fig:crimePolicies}.
Figure~\ref{fig:piSC} shows how the crime application interacts with its \textit{business collaborators} (external services and entities).
The interaction occurs by means of function calls (denoted by dotted lines in the figure).
Figure~\ref{fig:crimePolicies} shows the definition of three \textit{policies}, which define rules for service execution.
In our case we have policies for \textit{Security}, \textit{Performance} and \textit{Persistence}.

\paragraph*{$\pi$\textit{-PEWS} Model:}
this model is produced by a model-to-text transformation that takes a $\pi$-\textit{ServiceComposition} model and generates $\pi$-\textit{PEWS} specification code.
This code is a service composition program that can be compiled into an executable composition.
$\pi$-\textit{PEWS} models are expressed in a variant of the PEWS composition language.
\begin{figure}[t]
\begin{small}
%\begin{lstlisting}[label=list:rulePre,caption=ATL - piServiceComposition2piPEWS : Pre-condition Rule. ]
%\begin{lstlisting}[label=list:rulePre]
\begin{verbatim}
//Namespaces specify service URI
1 namespace twitter = www.twitter.com/service.wsdl
2 namespace googlemaps = maps.googleapis.com/maps/api/service
3 namespace amazondynamodb = rds.amazonaws.com/doc/2010-07-28/AmazonRDSv4.wsdl
4 namespace police = www.police.fr/service.wsdl
//Operations 
5 alias publishTwitter = portType/publishTwitter in twitter
6 alias searchCrime = portType/searchCrime in amazondynamodb 
7 alias showMap = portType/showMap in googlemaps 
//Services
8  service notifyCrime = publishCrime  . publishTwitter 
9  service trackCrime= searchCrime . verifyService
10 Service visualizeCrime =  showMap . getCrimeDetail
//Path
11 (notifyCrime.storeiCrime) || (trackCrime.visualizeCrime.getCrimeDetail)
//Contracts
12 defContract notifyCrimeContract{  isAppliedTo: notifyCrime
13	  requires: userId == ?? && passw == ?? && req(notifyCrime)  < 3
14 	      (OnFailureDo: NOT(action_publish(crime));
15	  ensures: publishTwitter(crime) == true  (OnFailureDo: skip); }
\end{verbatim}
% \end{lstlisting}
\end{small}
\caption{$\pi$-PEWS code for the crime tracking example (partial, simplified).\label{fig:pewscontract} }
\end{figure}
The $\pi$-PEWS program generated from the model in Figure~\ref{fig:policies} is partially presented in Figure~\ref{fig:pewscontract}. 
The figure shows a simplified program code, produced in accordance to the following guidelines:
\begin{itemize}
\item
Namespaces, identifying the addresses of external services are produced from the Business Collaborators of the higher-level model. 
We define four of them, corresponding to the Police, Twitter, Google Map and Amazon partners.
\item
Specific operations exported by each business collaborator are identified as an \textit{operation} of the program (to each operation is given an \texttt{alias}).
\item
The workflow in Figure~\ref{fig:piSC} is translated into the text at line 11.
\item
\textit{Contracts} are defined in $\pi$-PEWS as having pre-conditions (\texttt{requires}), post-conditions (\texttt{ensures}) and actions (\texttt{OnFailureDo}) to be executed whenever a condition is not verified. 
Contracts are generated from Policies (Figure~\ref{fig:crimePolicies}).
\end{itemize}