### 2013-10-22 SOFSEM2014/FinalVersion/main.tex
12 pages, Valeria's reccomndations.

\documentclass{llncs}
\usepackage{tikz}
\usepackage{amsfonts}
\usepackage{graphicx,url}
\usepackage[T1]{fontenc}
%\usepackage{psfig}
\usepackage{graphics}
\usepackage{amssymb,amsmath,latexsym}
\usepackage{listings} 
\usepackage{subfig}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{example}[theorem]{Example}
%\newtheorem{algorithm}[theorem]{Algorithm}
%\newtheorem{definition}[theorem]{Definition}

\usepackage[thmmarks]{ntheorem}

\usepackage{listings}
\theoremstyle{plain}
\theoremheaderfont{\normalfont\bfseries}
\theorembodyfont{\normalfont}
\theoremseparator{.}
\theoremindent0cm
\theoremnumbering{arabic}
\theoremsymbol{\ensuremath{\Box}} 
\newtheorem{defi}{Definition}%[section]

\theoremheaderfont{\itshape}
\newtheorem{examp}{Example}%[section]

\theoremstyle{plain}
\theoremheaderfont{\normalfont\bfseries}
\theorembodyfont{\normalfont}
\theoremseparator{}
\theoremindent0cm
\theoremnumbering{arabic}
%\theoremsymbol{\ensuremath{\Box}} 
\newtheorem{algo}{Algorithm}%[section]

\theoremstyle{plain} 
\theoremsymbol{\ensuremath{\Box}} 
\theoremseparator{.} 
\newtheorem{prop}{Property}

\lstset{numbers=right, numbersep=5pt, numberstyle=\tiny, stepnumber=1,escapechar=\!,columns=fullflexible,
        morekeywords={procedure,let,for,do,if,then,else,add,choose,end,while,
        true,false,rise,exception,extend,resume,to,return,function}}

\newcounter{numberInTrivlist}

\newenvironment{numtrivlist}{\begin{list}{\arabic{numberInTrivlist}.} 
                                         {\usecounter{numberInTrivlist}
                                          \setlength{\leftmargin}{0pt}
                                          \setlength{\rightmargin}{0pt}
                                          \setlength{\itemindent}{0pt}
                                          \setlength{\listparindent}{0pt}}}
                            {\end{list}}

\def\B{\hfill{$\Box$}}
\def\A{\hbox{${\cal A}$}}
\def\P{\hbox{${\cal P}$}}
\def\T{\hbox{${\cal T}$}}
\def\O{\hbox{${\cal O}$}}
\def\W{\hbox{${\cal W}$}}
\def\useOpt{\hbox{${has\_opt}$}}

\def\bigS{\hbox{${\cal S}$}}
\def\bigSd{\hbox{${\cal S}_D$}}
\def\bigSsm{\hbox{${\cal S}_D^{score}$}}

\newcommand{\ie}{\ifmmode \mathit{i.e.}\else \textit{i.e.}\fi}
\newcommand{\wrt}{\ifmmode \mathit{w.r.t.}\else \textit{w.r.t.}\fi}
\newcommand{\eg}{\ifmmode \mathit{e.g.}\else \textit{e.g.}\fi}
\newcommand{\tuple}[1]{\ifmmode \langle #1 \rangle\else $\langle #1 \rangle$\fi}

\usepackage{authblk}

\begin{document}
%
% --- Author Metadata here ---

% --- End of Author Metadata ---

\title{Supporting Non-Functional Requirements in Services Software Development Process: an MDD Approach}

\author{Valeria de Castro\inst{1} \and Martin A. Musicante\inst{2} \and Umberto Souza da Costa\inst{2} \and Pl\'acido~A.~de~Souza~Neto\inst{3} \and Genoveva Vargas-Solar\inst{4}}

\institute{Universidad Rey Juan Carlos -- M\'{o}stoles, Spain\\  \email{Valeria.deCastro@urjc.es} 
\and Federal University of Rio Grande do Norte (UFRN) -- Natal-RN, Brazil\\
\email{\{mam,umberto\}@dimap.ufrn.br}
\and Federal Technological Institute of Rio Grande do Norte (IFRN) -- Natal-RN, Brazil\\ \email{placido.neto@ifrn.edu.br}
\and French Council of Scientific Research (CNRS) -- Grenoble, France\\
\email{Genoveva.Vargas-Solar@imag.fr}
}

\maketitle
\begin{abstract}
This paper presents the $\pi$SOD-M method, an extension to the Service-Oriented Development Method (SOD-M) to support the development of services software by considering their functional and    non-functional requirements. 
Specifically, $\pi$SOD-M proposes: 
\textit{(i)} meta-models for representing non-functional requirements at different abstraction levels; 
\textit{(ii)} model-to-model transformation rules, useful to semi-automatically refine Platform Independent Models into Platform Specific Models; and 
\textit{(iii)} rules to transform Platform Specific Models into concrete implementations. 
In order to illustrate our proposal, the paper also describes how to apply the methodology to develop a proof of concept.
\end{abstract}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures,
%performance measures]

%\terms{Delphi theory}

\keywords{MDD, Service Oriented Applications, Non-functional Properties}

\section{Introduction}
\label{sec:intro}
\input{introduction}

\section{SOD-M}\label{sec:sodm}
\input{sodm}

\section{$\pi$SOD-M}\label{sec:pisodm}
\input{pisodm}

\section{Proof of Concept: \textit{Tracking Crimes}}\label{sec:poc}
\input{proofOfConcept}

\section{Related work}
\label{sec:related}
Over the last years, a number of approaches have been proposed for the
development of web services. 
These approaches range from the proposal of new 
languages for web service descriptions~\cite{bpel03,SBS04} 
to techniques to support phases of the development cycle of this kind of
software~\cite{lipari2007}. 
In general, these approaches concentrate on specific problems, like supporting transactions or
QoS, in order to improve the security and reliability of service-based
applications. 
Some proposals address service composition: workflow
definition~\cite{AalstHKB03,MuP06} or semantic equivalence between
services~\cite{BHM06}. 
%The proposed solutions come from many communities, 
%including those of Theoretical Computer Science~\cite{VA05,GGP08}, 
%Software Engineering~\cite{Aal03,MendesPDB09}, Programming
%Languages~\cite{MPC08,bpel03} and Databases~\cite{ABM01}.

Works dealing with non-functional properties in service-oriented 
development can be organized in two main groups: those working on the \textit{modeling 
of particular non-functional properties or QoS attributes} and those 
proposing \textit{architectures or frameworks to manage and 
validate QoS attributes in web service composition processes}. 
The first group considers specific non-functional concerns (e.g., security)
which is modelled and then associated to functional models of the application. 
The work of Chollet et al.~\cite{CholletL09} defines a proposal to associate non-functional 
quality properties (security properties) to functional activities in a web service composition model. Schmeling et al.~\cite{SchmelingCM11} present 
an approach and also a toolset for specifying and implementing non-functional 
concerns in web service compositions.
Non-functional concerns are modelled and then related to a service composition 
represented in a BPMN diagram. 
Ovaska et al.~\cite{OvaskaEHPA10} present an approach to 
support quality management at design time. 
Quality requirements are modelled in a first phase and then
represented in an architectural model where quality requirements are associated to some
components of the model. 
We propose to model functional and non-functional properties at the same time during the software process.
We claim what this approach simplifies the web service development task. 
Non-functional properties are represented in our work as constraints and policies 
but more general quality model such as the proposed by~\cite{Goeb2011,Klass2009} 
could be taken in consideration in further works.

The second group of works dealing with non-functional requirements for services 
propose specific architectures or frameworks to manage and validate 
QoS attributes in service composition processes~\cite{XiaoCZBOLH08,Babamir2010,Karunamurthy2012787}. 
Our proposal is similar to these, but focusses on more general non-functional properties.

Despite the variety of techniques proposed, there is not yet a consensus on a software methodology 
for web services. 
Some methodologies address the service-based 
development towards a standard or a new way to develop reliable applications. 
SOD-M and SOMF~\cite{somf} are MDD approaches for web services; 
S-Cube~\cite{scube2010book} is focused on the representation of business processes and 
service-based development; SOMA~\cite{soma} is a methodology 
for SOA solutions; 
%Extended SOA~\cite{PapazoglouH06} merges RUP and
%BPM~\cite{bpm} concepts for service modelling; 
DEVISE~\cite{DEVISE} is a
methodology for building service-based infrastructure for collaborative
enterprises. 
Other proposals include the WIED model~\cite{TongrungrojanaL04}, 
that acts as a bridge between business modeling and design models on one hand, and traditional 
approaches for software engineering applied to SOC on the other hand.

\section{Conclusions}
\label{sec:conclusions}
\input{conclusion}

\section*{Acknowledgements}
\small
This research is partly supported by the National Institute of Science and Technology for Software Engineering (INES\footnote{www.ines.org.br}), funded by CNPq (Brazil), grants 573964/2008-4 and 305619/2012-8; CAPES/UdelaR (Brazil/Uruguay) grant 021/ 2010; CAPES/STIC-AmSud (Brazil) grant 020/2010); MASAI project (TIN-2011-22617) financied by the Spanish Ministry of Science and Innovation and the Spanish Network on Service Science (TIN2011-15497-E) financed by the Spanish Ministry of Competitiveness and Economy. 
\normalsize

\bibliographystyle{splncs_srt}
\bibliography{biblio2} 
\end{document}