### 2014-09-10 SystematicMapping/background.tex
Systematic Mapping: outline

\subsection{Adding NFP to service compositions}

\subsection{Problem space and solution space}

\subsection{Models, methodologies and environments}

\subsection{Related work}