### 2014-09-10 SystematicMapping/background.tex

This section introduces the vocabulary and concepts related to nonfunctional properties of service based applications.

\subsection{Adding NFP to service compositions}

%\subsection{Problem space and solution space}

\subsection{Models, methodologies and environments}

\subsection{Related work}