### 2014-11-12 JournalToSubmit2014/figs/piPewsModel.tex

\begin{figure}[ht!]
\tiny
\centering

\begin{lstlisting}[label=list:flyingPigPiPEWS,caption=pi-PEWS
Specification: FlyingPig. ]

//Namespaces specify service URI
namespace orca = www.orca.mx/service.wsdl

//Operations
alias createCompanySpace = portType/createCompanySpace in orca
alias generateQuestionnaire = portType/generateQuestionnaire in orca
alias analizeAnswer = portType/analizeAnswer in orca
alias storeData = portType/storeData in orca
alias createReport = portType/createReport in orca
alias callForAmendments = portType/callForAmendments in orca

//Services
service receiveRequest(R, Id) = createCompanySpace(R, Id)

service generateNewInterface(Id, NULL) = ...

service createQuestionnaire(Id, Q) = generateQuestionnaire(Id, Q)

service notifyResponsable((Id, Q); NULL) = ...

service receiveAnswers((Id, T); P) =
								analizeAnswer((Id, T), NULL) 
								. storeData((Id, T), NULL)
								. (
								  (createReport(Id, P) . return(P))
								  + (callForAmendments(Id,T) . return(NULL))
								  )

//Path

receiveRequest(R, Id)
||
( generateNewInterface(Id, NULL)
  . createQuestionnaire(Id, Q)
  .  notifyResponsable((Id, Q); NULL
 )
||
 ( receiveAnswers((Id, T); P)
 . [P != NULL] STOP
 )*

//Policies

\end{lstlisting}
\label{fig:crimeInformationPEWS}  
\end{figure}