### 2014-11-12 JournalToSubmit2014/main.tex

\documentclass[twocolumn]{svjour3}

\usepackage{geometry}
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{amssymb} 
\usepackage{epstopdf}

\usepackage{amssymb}

%\usepackage{natbib}

%%% OUR MACROS %%%
\newcommand{\COMMENT}[1]{ }

\usepackage[usenames,dvipsnames]{xcolor}

\usepackage{amsmath}
%\usepackage[thmmarks,amsmath]{ntheorem}

\newcommand{\openbox}{\leavevmode
  \hbox to.77778em{%
  \hfil\vrule
  \vbox to.675em{\hrule width.6em\vfil\hrule}%
  \vrule\hfil}}

%\theoremstyle{plain}
%\theoremheaderfont{\normalfont\bfseries}
%\theorembodyfont{\normalfont}
%\theoremseparator{}
%\theoremindent0cm
%\theoremnumbering{arabic}
%\newtheorem{algo}{Algorithm}
%
%%\theoremstyle{plain}
%%%\theoremheaderfont{\normalfont\itshape}
%%\theoremheaderfont{\normalfont\bfseries}
%%\theorembodyfont{\normalfont}
%%\theoremseparator{}
%%\theoremindent0cm
%%\theoremnumbering{arabic}
%%\theoremsymbol{\ensuremath{\openbox}} 
%%\newtheorem{example}{Example}
%
%
%\theoremstyle{plain}
%\theoremheaderfont{\normalfont\bfseries}
%\theorembodyfont{\normalfont}
%\theoremseparator{.}
%\theoremindent0cm
%\theoremnumbering{arabic}
%\theoremsymbol{\ensuremath{\Box}} 
%\newtheorem{defi}{Definition}
%
%\theoremstyle{plain} 
%\theoremsymbol{\ensuremath{\Box}} 
%\theoremseparator{.} 
%\newtheorem{prop}{Property}

\def\FlyingPig{\textsl{FlyingPig}}

\newcounter{numberInTrivlist}

\newenvironment{numtrivlist}{\begin{list}{\rm \arabic{numberInTrivlist})} 
                                         {\usecounter{numberInTrivlist}
                                          \setlength{\leftmargin}{0pt}
                                          \setlength{\rightmargin}{0pt}
                                          \setlength{\itemindent}{12pt}
                                          \setlength{\listparindent}{0pt}}}
                            {\end{list}}

\newenvironment{itemizedTrivlist}{\begin{list}{\rm ~\hspace{2mm} $\bullet$\ } 
                                         {\setlength{\leftmargin}{0pt}
                                          \setlength{\rightmargin}{0pt}
                                          \setlength{\itemindent}{12pt}
                                          \setlength{\listparindent}{0pt}}}
                            {\end{list}}

\usepackage{listings}

\lstset{numbers=right, numbersep=5pt, numberstyle=\tiny, stepnumber=1,escapechar=\!,columns=fullflexible,
        morekeywords={procedure,let,for,do,if,then,else,add,choose,end,while,
        true,false,rise,exception,extend,resume,to,return,function}}

\newcommand{\pisodm}[0]{$\pi$SOD-M\xspace}

\begin{document}

\title{\pisodm: A New Methodology for Building Service-Oriented Applications}\thanks{\textsl{Contact author:}\\
Martin A. Musicante\\
UFRN - Cx Postal 1647, \\
Campus Universitário Lagoa Nova \\
59078-970, Natal, RN, Brazil\\[2mm]
Telephone: +55 84 3215 3814 ext 1.208\\
Mobile phone: +55 84 9653 4331\\
Fax: +55 84 3215 3781}

\author{Khalid Belhajjame \and Valeria~de~Castro \and Umberto~Souza~da~Costa \and Javier~A.~Espinosa-Oviedo \and Martin~A.~Musicante \and Pl\'acido~A.~Souza~Neto \and Genoveva~Vargas-Solar \and Jos\'e-Luis~Zechinelli-Martini}

\institute{K. Belhajjame
               \at Universit\'e de Paris - Dauphine -- Paris, France
               \and
               V. de Castro
               \at Universidad Rey Juan Carlos -- M\'{o}stoles, Spain
               \and
               U.S.Costa \and M.A. Musicante
               \at Universidade Federal do Rio Grande do Norte -- Natal, Brazil
               \and
               J.A. Espinosa-Oviedo, J-L Zechinelli-Martini
               \at Universidad de las Am\'ericas-Puebla, LAFMIA -- Cholula, Mexico
               \and 
               P. Souza Neto 
               \at Instituto Federal de Educa\c{c}\~{a}o, Ci\^{e}ncia e Tecnologia do Rio Grande do Norte -- Natal, Brazil
               \and
               G. Vargas-Solar
               \at CNRS, LIG-LAFMIA, Saint Martin d'H\`eres, France}

\maketitle

\begin{abstract}

Specifying non-functional requirements (NFRs) is a complex task, being usually dealt with on the later phases of the software process.
The late inclusion of NFRs in the development may compromise the quality of the deployed application.
This paper presents \pisodm, a methodology and associated tools that 
\textit{(i)}  allows the early specification of non-functional requirements in a principled way: users are abstracted away from low level details; 
\textit{(ii)} embraces the MDA philosophy, generating models (code) whenever possible.
The pro\-po\-sed solution has been utilized in the context of an industrial case study.
\end{abstract}

\keywords{MDA \and Non-Functional Requirements \and Service-based software process}

%% main text
%*********************************************************************************************************
\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Related Works}
\label{sec:relworks}
\input{relworks}

\section{Modeling reliable service compositions with \pisodm}\label{sec:motivation}
\input{modelingWithPISODM}

\section{Transformation rules}\label{sec:mmrules}
\input{transformationRules}

\section{Applying \pisodm: The \FlyingPig\ use case}
\label{sec:flyingPig}

\input{modelingFlyingPig}

\section{Conclusions and future work}\label{sec:conclusions}
\input{conclusions}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}