### 2015-03-31 JournalToSubmit2015/main.tex

\documentclass{ws-ijseke}
\usepackage{xspace}

\newcommand{\pisodm}[0]{$\pi$SOD-M\xspace}
\def\FlyingPig{\textsl{FlyingPig}}

\newcounter{numberInTrivlist}

\newenvironment{numtrivlist}{\begin{list}{\rm \arabic{numberInTrivlist})}
                                         {\usecounter{numberInTrivlist}
                                          \setlength{\leftmargin}{0pt}
                                          \setlength{\rightmargin}{0pt}
                                          \setlength{\itemindent}{12pt}
                                          \setlength{\listparindent}{0pt}}}
                            {\end{list}}

\newenvironment{itemizedTrivlist}{\begin{list}{\rm ~\hspace{2mm} $\bullet$\ }
                                         {\setlength{\leftmargin}{0pt}
                                          \setlength{\rightmargin}{0pt}
                                          \setlength{\itemindent}{12pt}
                                          \setlength{\listparindent}{0pt}}}
                            {\end{list}}

\begin{document}

\markboth{Authors' Names}
{Instructions for Typing Manuscripts (Paper's Title)}

%%%%%%%%%%%%%%%%%%%%% Publisher's Area please ignore %%%%%%%%%%%%%%%
%
\catchline{}{}{}{}{}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\pisodm: Building Service-Oriented Applications in the Presence of Non-Functional Requirements}

\author{Khalid Belhajjame}
\address{Universit\'e de Paris - Dauphine -- Paris, France\\
\email{kbelhajj@googlemail.com}
}

\author{Valeria~de~Castro} 
\address{Universidad Rey Juan Carlos -- M\'{o}stoles, Spain\\
\email{valeria.decastro@urjc.es}
}

\author{Umberto~Souza~da~Costa} 
\address{Universidade Federal do Rio Grande do Norte -- Natal, Brazil\\
\email{umberto@dimap.ufrn.br}
}

\author{Javier~A.~Espinosa-Oviedo} 
\address{Universidad de las Am\'ericas-Puebla, LAFMIA -- Cholula, Mexico\\
\email{javiera.espinosa@gmail.com}
}

\author{Martin~A.~Musicante} 
\address{Universidade Federal do Rio Grande do Norte -- Natal, Brazil\\
\email{mam@dimap.ufrn.br}
}

\author{Pl\'acido~A.~Souza~Neto} 
\address{Instituto Federal de Educa\c{c}\~{a}o, Ci\^{e}ncia e Tecnologia do Rio Grande do Norte -- Natal, Brazil\\
\email{placido.neto@ifrn.edu.br}
}

\author{Genoveva~Vargas-Solar} 
\address{CNRS, LIG-LAFMIA, Saint Martin d'H\`eres, France\\
\email{genoveva.vargas@imag.fr}
}

\author{Jos\'e-Luis~Zechinelli-Martini}
\address{Universidad de las Am\'ericas-Puebla, LAFMIA -- Cholula, Mexico\\
\email{joseluis.zechinelli@udlap.mx}
}

\maketitle

\begin{history}
\received{(Day Month Year)}
\revised{(Day Month Year)}
\accepted{(Day Month Year)}
%\comby{(xxxxxxxxxx)}
\end{history}

\begin{abstract}
Specifying non-functional requirements (NFRs) is a complex task, being usually dealt with on the later phases of the software process.
The late inclusion of NFRs in the development may compromise the quality of the deployed application.
This paper presents \pisodm, a method and associated tools that
\textit{(i)}  allows the early specification of non-functional requirements in a principled way: users are abstracted away from low level details;
\textit{(ii)} embraces the MDA philosophy, generating models (code) whenever possible.
The pro\-po\-sed solution has been utilized in the context of an industrial and real case study.
\end{abstract}

\keywords{MDA \and Non-Functional Requirements \and Service-based software process}

%% main text
%*********************************************************************************************************
\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Related Works}
\label{sec:relworks}
\input{relworks}

\section{Modeling reliable service compositions with \pisodm}\label{sec:motivation}
\input{modelingWithPISODM}

\section{Transformation rules}\label{sec:mmrules}
\input{transformationRules}

\section{Applying \pisodm: The \FlyingPig\ case study}
\label{sec:flyingPig}

\input{modelingFlyingPig}

\section{Conclusions and future work}\label{sec:conclusions}
\input{conclusions}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}